Post-Disturbance Diversity Trajectories in a Neotropical Forest

### 2018-01-30 AppendixI.Rmd
Manuscript update

---
title: "Post-Disturbance Diversity Trajectories in a Neotropical Forest"
author:
  - name: Ariane MIRABEL
    authsuperscript: 1*
  - name: Eric Marcon
    authsuperscript: 1
  - name: Bruno Hérault
    authsuperscript: 2
affiliation:
  - affsuperscript: 1
    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
    address: >
      Campus Agronomique,
      97310 Kourou,
      France.
  - affsuperscript: 2
    dptuniv: Department of Ecology, University of Edimburgh
    address: >
      Street address,
      Zip code,
      Country.
corrauthor:
    email: ariane.mirabel@ecofog.gf
    url: http://www.ecofog.gf/spip.php?article47
abstract: >   #Pour avoir plusieurs lignes qui suivent. Commencer à 2 caractères
  Résumé de l'article.
keywords: [mot-clés, séparés par des virgules]
journalinfo: Hal 00679993
archive: DOI xxxx
date: "`r format(Sys.time(), '%d %B %Y')`"
preamble: >  # Pour coordonne les packages LateX (supplémentaires)
  \usepackage{amsmath}
bibliography: references.bib # .bib à activer seulement pour git et word format (pandoc fait la biblio)
csl: 
lang: english # french
keywordlabel: Keywords
corrauthorlabel: Corresponding author
output:
  bookdown::gitbook
#   bookdown::pdf_book:
#     base_format: EcoFoG::article
#     keep_tex: TRUE
#  bookdown::word_document2
#Mettre un code dans les figures pour indiquer le code d'où elle provient. ../ pour référer à des sous dossiers
---

<!-- Options de knitr et de R (librairies, ...) -->

```{r Options, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
# Installation des packages si nécessaire et chargement
Library <- function(Packages) {
  InstallAndLoad <- function(Package) {
    if (!Package %in% installed.packages()[, 1]) {install.packages(Package)}
    require(Package, character.only = TRUE)
  }
  invisible(sapply(Packages, InstallAndLoad))
}
# Ajouter les packages nécessaires ici
source("Scripts/DrawingTools.R")
source("Scripts/GithubPages.R")
```

# Introduction
The large areas covered with tropical forests worldwide hold crucial commercial, social and cultural value. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support nutrients, water and carbon cycles and ensure cultural and human well-being. 
The demand for forests products and the shrink of unmanaged areas increased the human pressure on forests and currently threaten the maintenance of their structure, composition and function. 
The increasing demand for forests products, land-use changes and other human disturbances along with substantial climatic changes make a crucial issue of understanding the response of forests dynamics to such disturbance regimes [@Gibson2011a; @Morales-Hidalgo2015]. 
This has been addressed from the point of view of biomass production and carbon storage which resilience proved subjected to identified cyclic dynamics and therefore gives hope for the maintenance of forest ecosystem and their services. 
The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014]. 
Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of their resilience. 
Disturbance alter the biotic and abiotic parameters of the ecosystem in modifying the light availability, the heat and water fluxes, the plants productivity, etc. 
Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH). 
The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional profile of the whole community.
Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
Disturbance survey would first assess species turnover and communities convergence through stands taxonomic composition and unveil environmental filters through functional composition that is based on the biological attributes of plants. 
The set of functional traits chosen for this functional analysis considers functional traits representative of species functional strategy, fitness, and ecology as it is the case for traits of the leaf or stem economics.
Composition analysis would be continued with diversity analysis assessing for the whole community the richness and evenness of the taxonomic and functional abundance distribution.   
Here we examined the long-term trajectory, in the sense of the evolution along time, of tropical forests taxonomic and functional composition and diversity. 
We studied the regular inventories over 30 years of twelve Amazonian forest plots, either untouched or disturbed with three disturbance intensity . 
The inventories were combined with one of the largest functional trait database for tropical trees. We aimed to (i) explicit the impact of disturbance on the long-term trajectory of tropical forest communities composition and diversity, (ii) question the long-term convergence of the composition of communities and (iii) determine their resilience or not, after disturbance.

# Material and Methods

### Study site 
Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March @Wagner2011. 
Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
The experiment corresponds to a network of twelve 6.25ha plots where three logging, thinning and fuelwood treatments on an intensity gradient.
Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
The logging correspond to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3). 
Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling \ref{@Blanc2009}.

Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.

----------------------------------------------------------------------------------------------------------------------
                      ***Timber***            ***Thinning***               ***Fuelwood***               *%AGB lost*
-------------     --------------------     -------------------------     ------------------------     ---------------
***Control***                                                                                              0

***T1***            DBH ≥ 50 cm,
                    commercial species,                                                                [12%-33%]
                    ~ 10 trees/ha

***T2***            DBH ≥ 50 cm,            DBH ≥ 40 cm,
                    commercial species,    non-valuable species,                                        [33%-56%]
                    ~ 10 trees/ha          ~ 30 trees/ha

***T2***            DBH ≥ 50 cm,            DBH ≥ 50 cm,                   40 cm ≤ DBH ≤ 50 cm
                    commercial species,    non-valuable species,           non-valuable species,        [33%-56%]
                    ~ 10 trees/ha          ~ 15 trees/ha                   ~ 15 trees/ha
----------------------------------------------------------------------------------------------------------------------

### Inventories protocol and dataset correction
The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
During the inventories, trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
The botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years.
This inventory protocol rasied some mehodological issues given the variability of botanical identification practices along time and among plots. 
Because vernacular name attributed to trees usually correspond to a variety of botanical species, some taxonomic uncertainty remains whenever a trustworthy botanical name cannot be given thereafter. 
This taxonomic uncertainty was accounted for in the composition and diversity metrics throught a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
The uncertainty propagation is done by the extrapolation of complete inventories at genus level from real incomplete inventories, on the basis vernacular names replacement. The replacement of vernacular names uses a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.

The functional metrics used a set of 6 functional traits values representing traits of the leaf and wood economics and life history traits.
The trait database encompassed 4 traits of the leaves: leaves thickness (*L_thick*), toughness (*L_tough*), total chlorophyll content (*LCC*) and specific leaf area (leaf area per unit dry mass; *SLA*), 2 traits of the stem: wood specific gravity (*SWG*) and bark thickness (*B_thick)* and two life-history traits: maximum specific height (*Hmax*) and the seed mass (*Smass*). 
Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
To account for intraspecific traits variability, trait values at species level were assessed in sampling one of the measured individuals for iteration rather than averaging at species level. 
Missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits. 
The dataset comprised 257 botanical species. Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process.

### Composition and diversity metrics

The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories. 
Plots evolution along time was reportedcomparatively to the inventories in 1989, 5 years after disturbance, which correspond to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees). The distancing relative to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
The euclidean distance in the two-dimensional ordination space was used to visualize the distancing to reference inventory along time.
The functional composition was besides completed by trajectory along time of the community weighted means (CWM) for the 8 leaf, stem and life history traits. 
To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models. 
The null model for taxonomic composition randomly shuffled individuals among plots while preserving the overall species abundance and plots’ tree density. 
The null model for functional composition produced random traits assemblages by permuting the set of functional traits values among species.

Plots taxonomic diversity corresponded to species Richness and the Hill number translation of Shannon and Simpson indices. 
These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q). 
This order q grasps the balance between richness and evenness in the community as it determines the weight of common species in the diversity metric, with an growing emphasis on dominant species as q increases. 
The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on the 6 functional traits specific values.

# Results

Over time, 104423 trees and 696 botanical species were recorded. 
The disturbance proved to significantly impact for more than 30 years the functional and taxonomic composition and diversity of the stands.

### Taxonomic and functional composition

*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis. 
Classifications were based on either abundance-based Bray-Curtis \autoref{Fig1} or incidence-based Jaccard dissimilarity (data not shown).

Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
*The distance from the initial condition was significantly time-dependent (P<0.01).* 
The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \autoref{Fig1}. 
From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \autoref{Fig1},\autoref{Fig2}. 
The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \autoref{Fig2}. 
*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*

```{r Fig1, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance. Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
load("DB/FunctionalComposition_forGraph");
load("DB/TaxoComposition_forGraphs")

TaxoCompo(MatrepTaxo)
FunCompo(MatrepFun)

```

```{r Fig2, out.width = '100%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition. Distance are abundance-based Bary-Curtis metric. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes. "}

TaxoDist(MatrepTaxo)
FunDist(MatrepFun)

```

*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \autoref{Fig3}.
**To be coming**

```{r Fig3, echo=FALSE,fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness). The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3). The average CWM per treatment are drawn (solid line) drawn along with the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
```

### Disturbance impact on diversity

*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig4}. 
*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*

```{r Fig4, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The average value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
load("DB/ReplacementTraj_ForGraphs")

TaxoTraj(CompleteTaxo)

```

The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \autoref{Fig4}. 
The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig4}. 
The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig4} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.

*Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \autoref{Fig5}. 30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity. 
*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis.

```{r Fig5, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
load("DB/FunctionalTraj_ForGraphs")

FunTraj(CompleteFun)
```

# Discussion
-	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
-	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).   Impacts depend on the disturbance intensity, but further parameters may be involved. 
-	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging. Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance. Same conclusions otherwise.
-	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy. Disturbance generates a functional shift of communities toward acquisitive strategies. Ecosystems function rely on abundant species: changes in stands functioning is then expected.  (Flores, 2006  for differential sapling success after disturbance).
-	If the recovery is not achieved 30 years after disturbance, it is on the way to. Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
-	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.

# Appendix I _ Scheme of the taxonomic uncertainty handle

>>>>

---
title: "Appendix I _ Scheme of the taxonomic uncertainty handle"
author: "Ariane"
date: "30 janvier 2018"
output: html_document
---

### 2018-01-30 Manuscript1.Rmd
html files update
Manuscript update

Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
----------------------------------------------------------------------------------------------------------------------
                      ***Timber***            ***Thinning***               ***Fuelwood***               *%AGB lost*
-------------     --------------------     -------------------------     ------------------------     ---------------
***Control***                                                                                              0

@@

```{r Fig1, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance. Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
load("DB/FunctionalComposition_forGraph");
load("DB/TaxoComposition_forGraphs")

@@

# Appendix I _ Scheme of the taxonomic uncertainty handle

![**Fig. A1 _ ** Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing).](BayesianSheme.png)

>>>>

Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.

----------------------------------------------------------------------------------------------------------------------
                      ***Timber***            ***Thinning***               ***Fuelwood***               *%AGB lost*
-------------     --------------------     -------------------------     ------------------------     ---------------
***Control***                                                                                              0

@@

```{r Fig1, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance. Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
load("DB/FunctionalComposition_forGraph")
load("DB/TaxoComposition_forGraphs")

@@

`r if (knitr:::is_html_output()) '# References {-}'`

---
bibliography: ["references.bib"]
biblio-style: "apalike"
link-citations: true
---
